The semantic drift of quotations in blogspace: a case study in short-term cultural evolution

### 2016-08-10 brains-protocol.tex
Corrections after Camille's manuscript and emails

\newtext{We rely on a text corpus made of quotations extracted from online blog posts, and focus on their evolution.}
\newtext{\tbn{Indeed}}
Quotations appeared to be a perfect candidate to propose a first measure of \newtext{automatic} cognitive bias in \newtext{cultural transmission}.
First, they are usually cleanly delimited by quotation marks which greatly facilitates their detection in text corpora.
Second, they stem from a unique original version, and are ideally traceable back to that version.
Third, and most importantly, their duplication should \emph{a priori} be highly faithful, apart from cases of cropping:
not only should transformations be of moderate magnitude, but when specific words are not perfectly duplicated, it is safe to assume that the variation is due to involuntary cognitive bias --- as writers may expect any casual reader to easily verify, and thus criticize, the fidelity to the original quotation.

@@

To keep the analysis \newtext{tractable}, we focused on quotation transformations consisting \tb{of} the \emph{substitution} of a word by another word (and only those cases) in order to unambiguously discuss single word replacements.
\newtext{This restriction also allows us to more reliably infer the information that is missing in our data set, as explained further down (see ``Substitution model'').}
To quantify \tb{these} substitutions we decided to associate a number of features to each word, the variation of which we can statistically study.

@@

\subsection{\tb{Corpus-based} utterances}

@@

\newtext{Then, the authors gathered quotations in a graph and connected each pair that differed by no more than one word or that shared at least ten consecutive words (they tested this procedure with a number of different parameters, see~\citealp{leskovec_meme-tracking_2009}, for more details).}
We find for example the following variation of the above quote:
\begin{quote}
``these operations are \textbf{intended} to address funding pressures over quarter-end.''
\end{quote}
\newtext{Next, they applied a community detection algorithm} to that quotation graph to detect aggregates of tightly connected, \tb{i.e.} sufficiently similar, groups of quotations~\citep[see again][for more details]{leskovec_meme-tracking_2009}.
This analysis yielded the final data we had access to, with a total of about \num{70000} sets of quotations; each of these sets ideally contains all variations of a same parent utterance, along with their respective publication URLs and timestamps \newtext{(since the procedure cannot be perfect, sets of quotations contain occasional rogue unrelated variations that should have been discarded or assigned to another set)}.

Manual inspection of this data set revealed that it contains a significant number of everyday language quotations (such as ``it was much better than I expected'', ``did that just happen'', as well as many simple expletive-based sentences).
Their presence is largely due to random variations around casual expressions, while we are interested in transformations of news-related quotes causally linked to an original, identifiable utterance.
To filter them out, we exclude quotes with less than 5 words or \newtext{whose occurrences span more} than 80 days (indicating causally unrelated occurrences), as well as quotes not written in English.
\newtext{Clusters that are emptied by this procedure are therefore excluded.}
If, after this screening, a cluster's occurrences still span more than 80 days (because of short-lived but unrelated quotes far apart in time), we also exclude it.
\newtext{We eventually keep 50,427 clusters (out of 71,568; \hbox{i.e.} 70.5\%), containing a total of 141,324 unique quotes (out of 310,457; \hbox{i.e.} 45.5\%) making up about 2.60m occurrences (out of 7.67m; \hbox{i.e.} 33.9\%).\footnote{
The significantly larger loss in occurrences indicates that, on average, the clusters we lose contain more occurrences than those we keep, which is to be expected for everyday language utterances.
}
Even if we lose some real event-related utterances which are present in clusters lasting more than 80 days (one such lost quote, for instance, is ``the city is tired of me and the organization and I have run our course together''), we check that our \tb{filtering} approach fulfills its goals by coding a random sub-sample of 100 clusters: 35 of them are rejected by the filter, with 15 false negatives (rejected clusters that should have been kept) and 9 false positives (clusters kept when they should have been rejected), giving a precision score of 0.862 and a recall score of 0.789.
Furthermore, all but one of the 9 false positives are left with a single non-rejected quote, meaning those clusters are ignored by our substitution analysis; this brings the effective precision of our filter to 0.982.}\footnote{
\newtext{A similar analysis was made for language detection, which is part of the cluster filtering:
out of 100 randomly sampled quotes, 17 are rejected because their detected language is not English, with no false positives and 6 false negatives, giving a precision score of 1 and a recall score of 0.933.
Of the 6 false negatives, 4 had less than 5 tokens and would have been excluded by the cluster filter anyway.}
}

@@

\newtext{To compute \tb{these} features} we relied on the free association (FA) norms collected by \citet{nelson_university_2004}, which record the words that come to mind when someone is presented with a given cue.
As \citet{nelson_university_2004} explain, ``free association response probabilities index the likelihood that one word can cue another word to come to mind with minimal contextual constraints in effect.''
\newtext{Similar to what \citet{griffiths_google_2007} did, we first considered the directed \newtext{weighted} network formed by association norms, where nodes are words and edges are directed from cue to target word, with a weight equal to the association strength (that is the probability of that target word being produced when this particular cue is presented).
This network is of particular interest since it lets us define features that reflect the associations driving false memories in word lists~\citep{deese_prediction_1959}, a phenomenon which may be involved in the transformation of quotations.}

@@

Let us give an example:
\newtext{say the quotation ``These accusations are false and \textbf{absurd}'' ($q$) appears in two different blogs on January 19, and the slightly different quotation ``These accusations are false and \textbf{incoherent}'' ($q'$) appears in another blog on the 20th of January.
The second occurrence of $q$ can safely be assumed to be a faithful copy of the first one the same day.
And since $q$ is fairly prominent when $q'$ first appears, we could assume that the author of $q'$ on the 20th based herself on $q$, as is shown with a dashed line in Fig.~\ref{fig:substitution-unmodelled}.
Now say a third version, ``These \textbf{allegations} are false and \textbf{incoherent}'' ($q''$) also appears once on January 19 and once on January 20 after $q'$.
\tb{Here,} $q$ and $q''$ differ by two substitutions, so we discard the possibility that one was written based on the other (\tb{see} below for further details).
$q''$ is only one substitution away from $q'$ however, so we could also consider the first occurrence of $q''$ as a potential source for $q'$ on the 20th.
Conversely, the occurrence of $q''$ on the 20th could be considered as a substitution from $q'$, or as a faithful copy from its initial occurrence on January 19.}
(Options shown in Fig.~\ref{fig:substitution-unmodelled}.)

@@

We identified four binary parameters that differentiate potential models, such that the resulting 16 combinations cover most of the reasonable answers to inference uncertainties.
The first two parameters define the preceding time bin from which authors could have drawn a source when producing a new occurrence:
(1) \textbf{bin positions}, which can be aligned to midnight (as in the model presented above) or kept sliding (for each occurrence, use a bin that ends precisely at that occurrence);
(2) \textbf{bin span}, which can be \tb{$\Delta t=1$ day} %24 hours 
(as in the model above) or can be extended to start at the very first occurrence in the quotation family.
The other two parameters \tb{configure} rules on the selection of source and destination quotes of a substitution:
(3) \textbf{candidate sources} can be restricted to the most frequent quotations in the preceding time bin (as in the model above), or not (in which case all quotations in the preceding bin are candidate sources);
(4) \textbf{candidate destinations} can be restricted to quotations that do not appear in the preceding bin, or without restriction (as in the model above).
A substitution model, then, is the given of a value for each of those parameters;
it considers valid all the substitutions (and only those) where the source and destination follow the rules set out by the parameters.
If a destination has substitutions from multiple sources we count a single effective substitution where, for each feature, the value for the effective source word is the average of the values of the candidate source words.

@@

Put shortly a model defines how many times, and under what source and destination conditions, quote occurrences can be counted as substitutions.
Fig.~\ref{fig:substitution-modelled} shows the inferences made by the four models that use \tb{discrete} %ly positioned 
bins spanning 1 day:
later occurrences of $q'$ and $q''$ are counted as substitutions in Fig.~\ref{fig:substitution-modelled-majority-all} and Fig.~\ref{fig:substitution-modelled-all-all}, whereas in Fig.~\ref{fig:substitution-modelled-majority-exclude_past} and Fig.~\ref{fig:substitution-modelled-all-exclude_past} they are not.

@@

In practice \newtext{for the model first introduced above}, from the \newtext{2.60m} initial occurrences spread \tb{over} 50,427 quotation families, with significant redundancy (many quotes are indeed simple duplicates), we mine \newtext{40,868} substitutions.
From these substitutions we remove those featuring stop words, minor spelling changes (\hbox{e.g.} center/centre, November/Nov, Senator/Sen), abbreviations, spelled out numbers, \newtext{words unknown to WordNet, and deletions in substrings (which can appear as substitutions of non-deleted words)};
this eventually yields \newtext{6,318} valid substitutions\footnote{
\newtext{Manually coding a random subset of 100 substitutions to evaluate this last filter showed that 84 were true negatives, 5 were false positives, and 11 true positives, giving a recall score of .688.
Precision was evaluated over a random subset of 100 \emph{kept} substitutions, showing a score of .87.
Finally, note that excluding minor spelling changes does not bias our use of orthographic neighborhood density as a feature:
out of the first 100 substitutions coded for recall, those with Levenshtein distance equal to 1 (which is what orthographic neighborhood density codes,~\citealp{marian_clearpond:_2012}) were all typos or UK/US spelling changes, neither of which are relevant for this study.}} \newtext{\tbn{(before merging substitutions that share the same destination)}}.

@@

\tb{Substitutions usually} replace a word with another semantically related word: the new word achieves a similar meaning in the context of its sentence, while still slightly changing the implications or the attitude expressed by the author. The following examples illustrate this \tb{phenomenon}:

\begin{APAitemize}
	\item ``This is \{socialism~$\rightarrow$~welfare\} for the rich'',
	\item {[}The{]} ``perverse logic of \{clashes~$\rightarrow$~confrontation\} and violence'',
	\item ``This \{crisis~$\rightarrow$~problem\} did not develop overnight and it will not be solved overnight''.
\end{APAitemize}

\tb{However, the substituted and subtituting words are not so often direct synonyms:} only a third of all substitutions travel less than 3 hops on the hyponym-hypernym network defined by WordNet (direct synonyms count as 0~hops on this network), meaning that at least two thirds involve non-synonyms.\footnote{
\newtext{A similar \tb{phenomenon} is observed on the FA network, where about 104 clusters have substitutions traveling only 1 hop, 110 traveling 2 hops, 137 traveling 3, 72 traveling 4, and 13 traveling 5.
}}

@@

\newtext{Our question concerns the low-level properties of these substitutions: \tb{we ask (i) which words are targets of the substitutions and (ii) what change they are subjected to. To this end,}} we build the following two observables for each word feature.
First, we measure \newtext{which word features are more or less substituted compared to how often they would be if the process were random, in order to capture} the susceptibility for words to be the target of a substitution in a quote.
Second, we measure the change in word feature upon substitution, looking at the variation of a given feature between start and arrival words.
\newtext{Since sentence context is also central to this process, we extend these two observables by applying them to feature values relative to the distribution of feature values of the sentence in which a word appears.}

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{images/computed-figures/all-susceptibilities-pos.png}
    \caption{\newtext{\textbf{\tb{POS-related results:}} categories are simplified from the TreeTagger tag set: \emph{J} means adjective, \emph{N}~noun, \emph{R}~adverb, \emph{V}~verb, and \emph{Stopword-like} gathers together all the categories partially or completely affected by the stopword filter (see main text for details).
    The top panel shows the actual $s_{POS}$ and $s_{POS}^0$ counts. The bottom panel shows \tb{the substitution susceptibility $\sigma_{POS}$, which is the ratio between the two previous counts}.
    Confidence intervals are computed with the Goodman method for multinomial proportions.}
    }
    \label{fig:pos-susceptibilities}
\end{figure}

@@

We say that a word is \emph{substitutable} if it appears in a quote which undergoes a substitution, whether the substitution operates on that word or on another one.
\newtext{For a given group of words $g$, say all nouns, or all words in a particular range of values for a feature (\hbox{e.g.} words 2 to 4 letters long), susceptibility is computed as the ratio of $s_g$, the number of times words of that group are substituted, to $s_g^0$, the number of times words of that group would be substituted if substitutions fell randomly on substitutable words.
That is:}\footnote{
\newtext{$s_g^0$ is computed by summing, over all quotes undergoing a substitution, the ratio of the number of words in a quote that are from group $g$ to the number of words in the same quote that could have been the substitution's target.
If, for instance, half the content words of a given quote are nouns, such a quote contributes .5 to the total $s_{nouns}^0$.} \tb{Further, t}o avoid possible autocorrelation effects due to substitutions belonging to the same cluster (which are likely not statistically independent and may lead to overly optimistic confidence intervals), we scale $s_g$ and $s_g^0$ to count one for each cluster.
That is, each individual substitution contributes to $s_g$ or $s_g^0$ by only 1 over the number of substitutions in its cluster.}

@@

In other words, susceptibility measures \newtext{how much more or less a group of words $g$ actually gets substituted compared to picking targets at random in quotes undergoing substitutions}.
\newtext{By applying this measure to Part-of-Speech (POS) categories and feature bins (\hbox{e.g.} for a feature $\phi$ and a bin $[a, b]$, $g = \{w | \phi(w) \in [a, b]\}$), susceptibility measures the bias in} the selection of start words involved in substitutions, 
\tb{\hbox{i.e.} it measures the preferential selection of some word properties given a substitution.}
%\newtext{that is} the effect of word \newtext{properties} at the moment preceding the substitution when it is not yet known which word in the quotation will be substituted.

@@

Fig.~\ref{fig:pos-susceptibilities} gathers the results for POS groups.
A Goodman-based multinomial goodness-of-fit test~\citep{goodman_simultaneous_1965} shows that these categories have a significant effect on susceptibility ($p < .05$ in all substitution models), but this seems mostly due to the \emph{Stopword-like}\footnote{
\newtext{The \emph{Stopword-like} category gathers all the POS groups partially of completely affected by the stopword filter for substitutions: coordinating conjunctions, foreign words, prepositions, subordinating conjunctions, modals, possessive endings, punctuation symbols and interjections.
Note that the stopword filter equally affects $s$ and $s^0$, so while both numbers are very small for these categories (which led us to group them together) the reported susceptibility is not biased by this filtering.}
}
and \emph{Adverb} categories.
Indeed, detailing which categories are out of their confidence region under $\mathcal{H}_0$ shows that susceptibility for stopword-likes is significantly below 1 (confirmed in all substitution models), as is that for adverbs in 13 of the 16 substitution models;
none of the other categories are significantly different from $\mathcal{H}_0$ (except nouns which appear significantly above 1 in a single substitution model).
While we acknowledge the low susceptibilities of adverbs and stopword-likes, these categories concern less than 7\% of all substitutions under $\mathcal{H}_0$ (and even less in the actual data);
it seems, then, that POS categories do not capture any strong bias in the selection of substitution targets.

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{images/computed-figures/paper-variations_regression-globals_to_globalsvariation-heatmap.png}
    \caption{\newtext{\textbf{Feature variations regression coefficients}:
    source feature values \tb{(columns)} and feature variations \tb{(rows)} were normalized to $[0; 1]$ to ensure the coefficients are comparable.
    Significance levels for individual \emph{t}-tests against the hypothesis of a null coefficient are denoted by stars below the corresponding coefficient (*** for $p \leq .001$, ** for $p \leq .01$, * for $p \leq .05$, and nothing when $p > .05$).
	Frequency has a slight effect on Age of acquisition and Clustering coefficient, with small coefficients compared to the respective diagonal ones.
	Aside from those two, only diagonal values are significantly non null.
	}}
    \label{fig:feature-variations_regression-global}
\end{figure}

>>>>

\newtext{We rely on a text corpus made of quotations extracted from online blog posts, and focus on their evolution.}
Quotations appeared to be a perfect candidate to propose a first measure of \newtext{automatic} cognitive bias in \newtext{cultural transmission}.
First, they are usually cleanly delimited by quotation marks which greatly facilitates their detection in text corpora.
Second, they stem from a unique original version, and are ideally traceable back to that version.
Third, and most importantly, their duplication should \emph{a priori} be highly faithful, apart from cases of cropping:
not only should transformations be of moderate magnitude, but when specific words are not perfectly duplicated, it is safe to assume that the variation is due to involuntary cognitive bias --- as writers may expect any casual reader to easily verify, and thus criticize, the fidelity to the original quotation.

@@

To keep the analysis \newtext{tractable}, we focused on quotation transformations consisting of the \emph{substitution} of a word by another word (and only those cases) in order to unambiguously discuss single word replacements.
\newtext{This restriction also allows us to more reliably infer the information that is missing in our data set, as explained further down (see ``Substitution model'').}
To quantify these substitutions we decided to associate a number of features to each word, the variation of which we can statistically study.

@@

\subsection{Corpus-based utterances}

@@

\newtext{Then, the authors gathered quotations in a graph and connected each pair that differed by no more than one word or that shared at least ten consecutive words (they tested this procedure with a number of different parameters, see~\citealp{leskovec_meme-tracking_2009}, for more details).}
We find for example the following variation of the above quote:
\begin{quote}
``these operations are \textbf{intended} to address funding pressures over quarter-end.''
\end{quote}
\newtext{Next, they applied a community detection algorithm} to that quotation graph to detect aggregates of tightly connected, that is
% Le APA style guide force "that is" au lieu de "i.e." hors de parenthèses
sufficiently similar, groups of quotations~\citep[see again][for more details]{leskovec_meme-tracking_2009}.
This analysis yielded the final data we had access to, with a total of about \num{70000} sets of quotations; each of these sets ideally contains all variations of a same parent utterance, along with their respective publication URLs and timestamps \newtext{(since the procedure cannot be perfect, sets of quotations contain occasional rogue unrelated variations that should have been discarded or assigned to another set)}.

Manual inspection of this data set revealed that it contains a significant number of everyday language quotations (such as ``it was much better than I expected'', ``did that just happen'', as well as many simple expletive-based sentences).
Their presence is largely due to random variations around casual expressions, while we are interested in transformations of news-related quotes causally linked to an original, identifiable utterance.
To filter them out, we exclude quotes with less than 5 words or \newtext{whose occurrences span more} than 80 days (indicating causally unrelated occurrences), as well as quotes not written in English.
\newtext{Clusters that are emptied by this procedure are therefore excluded.}
If, after this screening, a cluster's occurrences still span more than 80 days (because of short-lived but unrelated quotes far apart in time), we also exclude it.
\newtext{We eventually keep 50,427 clusters (out of 71,568; \hbox{i.e.} 70.5\%), containing a total of 141,324 unique quotes (out of 310,457; \hbox{i.e.} 45.5\%) making up about 2.60m occurrences (out of 7.67m; \hbox{i.e.} 33.9\%).\footnote{
The significantly larger loss in occurrences indicates that, on average, the clusters we lose contain more occurrences than those we keep, which is to be expected for everyday language utterances.
}
Even if we lose some real event-related utterances which are present in clusters lasting more than 80 days (one such lost quote, for instance, is ``the city is tired of me and the organization and I have run our course together''), we check that our filtering approach fulfills its goals by coding a random sub-sample of 100 clusters: 35 of them are rejected by the filter, with 15 false negatives (rejected clusters that should have been kept) and 9 false positives (clusters kept when they should have been rejected), giving a precision score of 0.862 and a recall score of 0.789.
Furthermore, all but one of the 9 false positives are left with a single non-rejected quote, meaning those clusters are ignored by our substitution analysis; this brings the effective precision of our filter to 0.982.}\footnote{
\newtext{A similar analysis was made for language detection, which is part of the cluster filtering:
out of 100 randomly sampled quotes, 17 are rejected because their detected language is not English, with no false positives and 6 false negatives, giving a precision score of 1 and a recall score of 0.933.
Of the 6 false negatives, 4 had less than 5 tokens and would have been excluded by the cluster filter anyway.}
}

@@

\newtext{To compute these features} we relied on the free association (FA) norms collected by \citet{nelson_university_2004}, which record the words that come to mind when someone is presented with a given cue.
As \citet{nelson_university_2004} explain, ``free association response probabilities index the likelihood that one word can cue another word to come to mind with minimal contextual constraints in effect.''
\newtext{Similar to what \citet{griffiths_google_2007} did, we first considered the directed \newtext{weighted} network formed by association norms, where nodes are words and edges are directed from cue to target word, with a weight equal to the association strength (that is the probability of that target word being produced when this particular cue is presented).
This network is of particular interest since it lets us define features that reflect the associations driving false memories in word lists~\citep{deese_prediction_1959}, a phenomenon which may be involved in the transformation of quotations.}

@@

Let us give an example:
\newtext{say the quotation ``These accusations are false and \textbf{absurd}'' ($q$) appears in two different blogs on January 19, and the slightly different quotation ``These accusations are false and \textbf{incoherent}'' ($q'$) appears in another blog on the 20th of January.
The second occurrence of $q$ can safely be assumed to be a faithful copy of the first one the same day.
And since $q$ is fairly prominent when $q'$ first appears, we could assume that the author of $q'$ on the 20th based herself on $q$, as is shown with a dashed line in Fig.~\ref{fig:substitution-unmodelled}.
Now say a third version, ``These \textbf{allegations} are false and \textbf{incoherent}'' ($q''$) also appears once on January 19 and once on January 20 after $q'$.
Here, $q$ and $q''$ differ by two substitutions, so we discard the possibility that one was written based on the other (see below for further details).
$q''$ is only one substitution away from $q'$ however, so we could also consider the first occurrence of $q''$ as a potential source for $q'$ on the 20th.
Conversely, the occurrence of $q''$ on the 20th could be considered as a substitution from $q'$, or as a faithful copy from its initial occurrence on January 19.}
(Options shown in Fig.~\ref{fig:substitution-unmodelled}.)

@@

We identified four binary parameters that differentiate potential models, such that the resulting 16 combinations cover most of the reasonable answers to inference uncertainties.
The first two parameters define the preceding time bin from which authors could have drawn a source when producing a new occurrence:
(1) \textbf{bin positions}, which can be aligned to midnight (as in the model presented above) or kept sliding (for each occurrence, use a bin that ends precisely at that occurrence);
(2) \textbf{bin span}, which can be $\Delta t=1$ day %24 hours 
(as in the model above) or can be extended to start at the very first occurrence in the quotation family.
The other two parameters configure rules on the selection of source and destination quotes of a substitution:
(3) \textbf{candidate sources} can be restricted to the most frequent quotations in the preceding time bin (as in the model above), or not (in which case all quotations in the preceding bin are candidate sources);
(4) \textbf{candidate destinations} can be restricted to quotations that do not appear in the preceding bin, or without restriction (as in the model above).
A substitution model, then, is the given of a value for each of those parameters;
it considers valid all the substitutions (and only those) where the source and destination follow the rules set out by the parameters.
If a destination has substitutions from multiple sources we count a single effective substitution where, for each feature, the value for the effective source word is the average of the values of the candidate source words.

@@

Put shortly a model defines how many times, and under what source and destination conditions, quote occurrences can be counted as substitutions.
Fig.~\ref{fig:substitution-modelled} shows the inferences made by the four models that use discrete %ly positioned 
bins spanning 1 day:
later occurrences of $q'$ and $q''$ are counted as substitutions in Fig.~\ref{fig:substitution-modelled-majority-all} and Fig.~\ref{fig:substitution-modelled-all-all}, whereas in Fig.~\ref{fig:substitution-modelled-majority-exclude_past} and Fig.~\ref{fig:substitution-modelled-all-exclude_past} they are not.

@@

In practice \newtext{for the model first introduced above}, from the \newtext{2.60m} initial occurrences spread over 50,427 quotation families, with significant redundancy (many quotes are indeed simple duplicates), we mine \newtext{40,868} substitutions.
From these substitutions we remove those featuring stop words, minor spelling changes (\hbox{e.g.} center/centre, November/Nov, Senator/Sen), abbreviations, spelled out numbers, \newtext{words unknown to WordNet, and deletions in substrings (which can appear as substitutions of non-deleted words)};
this eventually yields \newtext{6,177} valid substitutions.\footnote{
\newtext{Manually coding a random subset of 100 substitutions to evaluate this last filter showed that 84 were true negatives, 5 were false positives, and 11 true positives, giving a recall score of .688.
Precision was evaluated over a random subset of 100 \emph{kept} substitutions, showing a score of .87.
Finally, note that excluding minor spelling changes does not bias our use of orthographic neighborhood density as a feature:
out of the first 100 substitutions coded for recall, those with Levenshtein distance equal to 1 (which is what orthographic neighborhood density codes,~\citealp{marian_clearpond:_2012}) were all typos or UK/US spelling changes, neither of which are relevant for this study.}}

@@

Substitutions usually replace a word with another semantically related word:
\tb{manual observation of a random subset of 100 substitutions shows that, compared to the word it replaces,} the new word often achieves a similar meaning in the context of its sentence while still slightly changing the implications or the attitude expressed by the author.
The following examples illustrate this phenomenon:\footnote{
\newtext{However, the substituted and substituting words are not so often direct synonyms:
only a third of all substitutions travel less than 3 hops on the hyponym-hypernym network defined by WordNet (direct synonyms count as 0~hops on this network), meaning that at least two thirds involve non-synonyms.
A similar phenomenon is observed on the FA network, where about 104 clusters have substitutions traveling only 1 hop, 110 traveling 2 hops, 137 traveling 3, 72 traveling 4, and 13 traveling 5.
}}
\begin{APAitemize}
	\item ``This is \{socialism~$\rightarrow$~welfare\} for the rich'',
	\item {[}The{]} ``perverse logic of \{clashes~$\rightarrow$~confrontation\} and violence'',
	\item ``This \{crisis~$\rightarrow$~problem\} did not develop overnight and it will not be solved overnight''.
\end{APAitemize}

@@

\newtext{Our question concerns the low-level properties of these substitutions: we ask (i) which words are targets of the substitutions and (ii) what change they are subjected to. To this end,} we build the following two observables for each word feature.
First, we measure \newtext{which word features are more or less substituted compared to how often they would be if the process were random, in order to capture} the susceptibility for words to be the target of a substitution in a quote.
Second, we measure the change in word feature upon substitution, looking at the variation of a given feature between start and arrival words.
\newtext{Since sentence context is also central to this process, we extend these two observables by applying them to feature values relative to the distribution of feature values of the sentence in which a word appears.}

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{images/computed-figures/all-susceptibilities-pos.png}
    \caption{\newtext{\textbf{POS-related results:}
    categories are simplified from the TreeTagger tag set: \emph{J} means adjective, \emph{N}~noun, \emph{R}~adverb, \emph{V}~verb, and \emph{Stopword-like} gathers together all the categories \tb{essentially made of stopwords} (see main text for details).
    The top panel shows the actual $s_{POS}$ and $s_{POS}^0$ counts. The bottom panel shows the substitution susceptibility $\sigma_{POS}$, which is the ratio between the two previous counts.
    Confidence intervals are computed with the Goodman method for multinomial proportions.}
    }
    \label{fig:pos-susceptibilities}
\end{figure}

@@

We say that a word is \emph{substitutable} if it appears in a quote which undergoes a substitution, whether the substitution operates on that word or on another one.
\newtext{For a given group of words $g$, say all nouns, or all words in a particular range of values for a feature (\hbox{e.g.} words 2 to 4 letters long), susceptibility is computed as the ratio of $s_g$, the number of times words of that group are substituted, to $s_g^0$, the number of times words of that group would be substituted if substitutions fell randomly on substitutable words.
That is:}\footnote{
\newtext{$s_g^0$ is computed by summing, over all quotes undergoing a substitution, the ratio of the number of words in a quote that are from group $g$ to the number of words in the same quote that could have been the substitution's target.
If, for instance, half the content words of a given quote are nouns, such a quote contributes .5 to the total $s_{nouns}^0$.
Further, to avoid possible autocorrelation effects due to substitutions belonging to the same cluster (which are likely not statistically independent and may lead to overly optimistic confidence intervals), we scale $s_g$ and $s_g^0$ to count one for each cluster.
\tb{That is, each quote cluster has a maximum contribution of 1, computed as the average contribution of all substitutions in that cluster.}
}}

@@

In other words, susceptibility measures \newtext{how much more or less a group of words $g$ actually gets substituted compared to picking targets at random in quotes undergoing substitutions}.
\newtext{By applying this measure to Part-of-Speech (POS) categories and feature bins (\hbox{e.g.} for a feature $\phi$ and a bin $[a, b]$, $g = \{w | \phi(w) \in [a, b]\}$), susceptibility measures the bias in} the selection of start words involved in substitutions, 
\hbox{i.e.} \newtext{it measures the preferential selection of some word properties \tb{for} substitution.}
%\newtext{that is} the effect of word \newtext{properties} at the moment preceding the substitution when it is not yet known which word in the quotation will be substituted.

@@

Fig.~\ref{fig:pos-susceptibilities} gathers the results for POS groups.
A Goodman-based multinomial goodness-of-fit test~\citep{goodman_simultaneous_1965} shows that these categories have a significant effect on susceptibility ($p < .05$ in all substitution models), but this seems mostly due to the \emph{Stopword-like}\footnote{
\newtext{The \emph{Stopword-like} category gathers all the POS groups \tb{essentially made of stopwords (coordinating conjunctions, foreign words, prepositions, subordinating conjunctions, modals, possessive endings, punctuation symbols and interjections), which are affected by our decision to exclude stopwords from the analysis.
Indeed substitutions falling on stopwords are filtered out (affecting $s$), and stopwords are never counted as substitutable in the computation of $s^0$.
Both numbers are subsequently very small for these categories (which led us to group them together) but the reported susceptibility is left unbiased.}}
}
and \emph{Adverb} categories.
Indeed, detailing which categories are out of their confidence region under $\mathcal{H}_0$ shows that susceptibility for stopword-likes is significantly below 1 (confirmed in all substitution models), as is that for adverbs in 13 of the 16 substitution models;
none of the other categories are significantly different from $\mathcal{H}_0$ (except nouns which appear significantly above 1 in a single substitution model).
While we acknowledge the low susceptibilities of adverbs and stopword-likes, these categories concern less than 7\% of all substitutions under $\mathcal{H}_0$ (and even less in the actual data);
it seems, then, that POS categories do not capture any strong bias in the selection of substitution targets.

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{images/computed-figures/paper-variations_regression-globals_to_globalsvariation-heatmap.png}
    \caption{\newtext{\textbf{Feature variations regression coefficients}:
    source feature values (columns) and feature variations (rows) were normalized to $[0; 1]$ to ensure the coefficients are comparable.
    Significance levels for individual \emph{t}-tests against the hypothesis of a null coefficient are denoted by stars below the corresponding coefficient (*** for $p \leq .001$, ** for $p \leq .01$, * for $p \leq .05$, and nothing when $p > .05$).
	Frequency has a slight effect on Age of acquisition and Clustering coefficient, with small coefficients compared to the respective diagonal ones.
	Aside from those two, only diagonal values are significantly non null.
	}}
    \label{fig:feature-variations_regression-global}
\end{figure}

### 2016-08-10 brainscopycut.tex
Corrections after Camille's manuscript and emails

\title{The semantic drift of quotations in blogspace: a case study in short-term cultural evolution}
\shorttitle{The semantic drift of quotations in blogspace}
\date{}
\twoauthors{Sébastien Lerique}{Camille Roth}
\twoaffiliations{Centre d'Analyse et de Mathématique Sociales, UMR 8557 CNRS/EHESS\\190 av. de France, F-75013 Paris\\and Centre Marc Bloch Berlin, UMIFRE 14 CNRS/MAEE/HU\\Friedrichstr. 191, D-10117 Berlin}{CNRS\\Centre Marc Bloch Berlin, UMIFRE 14 CNRS/MAEE/HU\\Friedrichstr. 191, D-10117 Berlin}
\abstract{
\newtext{We present a case study which connects psycholinguistics with the field of cultural evolution, in order to test for the existence of cultural attractors in the evolution of quotations.
Such attractors have been \tb{proposed} as \tb{a concept useful for} understanding cultural evolution %and its reciprocal influences 
\tb{in relation} with \tb{individual} cognition, but their existence has been hard to test.
\tb{We empirically} focus on \tb{the online transformation of} quotations when they are copied from blog to blog or media website: by coding words with a number of well-studied lexical features, we show that the way words are substituted in quotations is consistent (1) with the hypothesis of cultural attractors, and (2) with known effects of the word features.
In particular, words known to be harder to recall in lists have a higher tendency to be substituted, and words easier to recall are produced instead.
Our results support the hypothesis that cultural attractors can \tb{result} from the combination of individual cognitive biases in the interpretation and reproduction of representations.}
}
\authornote{Correspondence should be directed to lerique@cmb.hu-berlin.de and roth@cmb.hu-berlin.de}
\keywords{\newtext{word recall}; recollection bias; semantic network; cultural evolution; cultural attraction; data mining; big data; psycholinguistics}
%\note{\TB{Note XXXX (version? date?)}}

>>>>

\title{The semantic drift of quotations in blogspace: a case study in short-term cultural evolution}
\shorttitle{The semantic drift of quotations in blogspace}
\date{}
\twoauthors{Sébastien Lerique}{Camille Roth}
\twoaffiliations{Centre d'Analyse et de Mathématique Sociales, UMR 8557 CNRS/EHESS\\190 av. de France, F-75013 Paris\\and Centre Marc Bloch Berlin, UMIFRE 14 CNRS/MAEE/HU\\Friedrichstr. 191, D-10117 Berlin}{CNRS\\Centre Marc Bloch Berlin, UMIFRE 14 CNRS/MAEE/HU\\Friedrichstr. 191, D-10117 Berlin}
\abstract{
\newtext{We present an empirical case study which connects psycholinguistics with the field of cultural evolution, in order to test for the existence of cultural attractors in the evolution of quotations.
Such attractors have been proposed as a useful concept for understanding cultural evolution in relation with individual cognition, but their existence has been hard to test.
We focus on the transformation of quotations when they are copied from blog to blog or media website:
by coding words with a number of well-studied lexical features, we show that the way words are substituted in quotations is consistent (1) with the hypothesis of cultural attractors, and (2) with known effects of the word features.
In particular, words known to be harder to recall in lists have a higher tendency to be substituted, and words easier to recall are produced instead.
Our results support the hypothesis that cultural attractors can result from the combination of individual cognitive biases in the interpretation and reproduction of representations.}
}
\authornote{Correspondence should be directed to lerique@cmb.hu-berlin.de and roth@cmb.hu-berlin.de}
\keywords{\newtext{word recall}; recollection bias; semantic network; cultural evolution; cultural attraction; data mining; big data; psycholinguistics}
%\note{\TB{Note XXXX (version? date?)}}
